Data preparation

### 2025-07-29 procesamiento/02-prod_prep-country.Rmd
data for reproducibility

---
title: "Data preparation"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
output: 
  html_document: 
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
      number_sections: yes
    code_folding: show  
    number_sections: yes
editor_options: 
  chunk_output_type: console
---

@@

```{r setup}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      echo = TRUE 
                      )
options(scipen=9999) # desactivar notacion cientifica
if (!require("pacman")) install.packages("pacman") # instalar pacman
load(here::here("input/data/proc/study1.RData"))
pacman::p_load(dplyr,sjmisc,sjPlot,sjlabelled,tidyverse,haven,MLMusingR)
```

@@

Revisar:

https://wid.world/codes-dictionary#post-tax-income

@@

**Otero et.al (2023)**

```{r otero23}
otero <- 
  read.csv(file = here::here("input/data/original/otero23.csv"),sep = ";")
otero$ccode <- countrycode::countrycode(sourcevar = otero$iso3n,origin = "iso2c",destination = "iso3n")
otero$gdp <- gsub('\\,', '',otero$gdp) 
otero<- otero %>% select(ccode,gv_spen,individualism)
```

@@

ilo_df <- left_join(long_ilo_tax,long_ilo_spe) %>% filter(year==2016)

# long_ilo_spe$country <- countrycode::countrycode(sourcevar = long_ilo_spe$ccode,origin = "iso3n",destination = "iso3c")

# Plot points connected by a line representing time
# ggplot(long_ilo_spe, aes(x = factor(year), y = govspe, group = factor(country), color = factor(country))) +
#   geom_line() +
#   geom_point(size = 3) +
#   labs(title = "Points Connected by a Line Representing Time", x = "Year", y = "Value") +
#   # scale_x_discrete(labels = c("2016", "2017", "2018")) +
#   theme_minimal()
# 
# 
# # Plot points connected by a line representing time
# ggplot(long_ilo_tax, aes(x = factor(year), y = taxrev, group = factor(country), color = factor(country))) +
#   geom_line() +
#   geom_point(size = 3) +
#   labs(title = "Points Connected by a Line Representing Time", x = "Year", y = "Value") +
#   # scale_x_discrete(labels = c("2016", "2017", "2018")) +
#   theme_minimal()

```

@@

```{r merge}
macro_data <- 
  left_join(wiid_2017,solt_2017) %>%
  left_join(pen_2017) %>% 
  left_join(otero) %>%
  left_join(ocde) %>% 
  left_join(wid_df1) %>% 
  left_join(ilo_df)

save(macro_data,file = here::here("input/data/proc/study1_macro_data.RData"))

df1$ccode <- as.numeric(df1$country)
df2 <- left_join(df1,macro_data,by="ccode")
df2 <- 
df2 %>% mutate(country2=countrycode::countrycode(sourcevar = ccode,origin = "iso3n",destination = "iso3c"),
               country3=countrycode::countrycode(sourcevar = ccode,origin = "iso3n",destination = "country.name"),
               region=countrycode::countrycode(sourcevar = ccode,origin = "iso3n",destination = "region23")) %>% 
  select(ccode,country2,country3,everything())
```

# Class inequality

```{r}
summary(df1$incomepcap)

df_class3_ineq <- 
df2 %>%   
  group_by(country2,dclass3res_V) %>%
  summarise_at(c("incomepcap","ednum","wid_gini_disp"), mean, na.rm = TRUE) %>% 
  as.data.frame() %>%
  na.omit() %>% 
  dplyr::select(country2,dclass3res_V,incomepcap,ednum,wid_gini_disp) %>% 
  tidyr::pivot_wider(names_from = "dclass3res_V", values_from = c("incomepcap","ednum")) %>% 
  mutate(ratio_class_inc = `incomepcap_Service Class (I+II)`/`incomepcap_Working Class (V+VI+VII)`,
         ratio_class_educ  = `ednum_Service Class (I+II)`/`ednum_Working Class (V+VI+VII)`) %>% 
  as.data.frame()

df_class3_ineq %>% select(wid_gini_disp,ratio_class_inc,ratio_class_educ) %>% tab_corr()

df_class6_ineq <- 
df2 %>%   
  group_by(country2,dclass6res) %>%
  summarise_at(c("incomepcap","ednum","wid_gini_disp"), mean, na.rm = TRUE) %>% 
  as.data.frame() %>%
  na.omit() %>% 
  dplyr::select(country2,dclass6res,incomepcap,ednum,wid_gini_disp) %>% 
  tidyr::pivot_wider(names_from = "dclass6res", values_from = c("incomepcap","ednum")) %>% 
  mutate(ratio_class_inc = `incomepcap_Upper Service class`/`incomepcap_Unskilled working class`,
         ratio_class_educ  = `ednum_Upper Service class`/`ednum_Unskilled working class`) %>% 
  as.data.frame() 
df_class_ineq %>% select(wid_gini_disp,ratio_class_inc,ratio_class_educ) %>% tab_corr()
```

```{r}

income_data

# Example data
# Assuming you have a data frame called 'income_data' with columns: country, social_class, and average_income

# Calculate all pairs of income ratios within each country
income_ratios <- income_data %>%
  group_by(country) %>%
  mutate(income_ratio = combn(average_income, 2, FUN = function(x) x[1] / x[2])) %>%
  ungroup()

# Print the result
print(income_ratios)

```

```{r save}
save(df2,file = here::here("input/data/proc/study1_country.RData"))
# haven::write_dta(data = df2,path = here::here("input/data/proc/study1_country.dta"))
sjPlot::view_df(df2,file =here::here("input/data/proc/codebook-s1.html") )
```

>>>>

---
title: "Data preparation - country"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
output: 
  html_document: 
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
      number_sections: yes
    code_folding: show  
    number_sections: yes
editor_options: 
  chunk_output_type: console
---

@@

```{r setup}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      echo = TRUE 
                      )
options(scipen=9999)
if (!require("pacman")) install.packages("pacman") 
load(here::here("input/data/proc/study1.RData"))
pacman::p_load(dplyr,sjmisc,sjPlot,sjlabelled,tidyverse,haven,MLMusingR)
```

@@

@@

@@

ilo_df <- left_join(long_ilo_tax,long_ilo_spe) %>% filter(year==2016)
```

@@

```{r merge}
macro_data <- 
  left_join(wiid_2017,solt_2017) %>%
  left_join(pen_2017) %>% 
  left_join(ocde) %>% 
  left_join(wid_df1) %>% 
  left_join(ilo_df)

save(macro_data,file = here::here("input/data/proc/study1_macro_data.RData"))
df1$ccode <- as.numeric(df1$country)
df2 <- left_join(df1,macro_data,by="ccode")
df2 <- 
df2 %>% mutate(country2=countrycode::countrycode(sourcevar = ccode,origin = "iso3n",destination = "iso3c"),
               country3=countrycode::countrycode(sourcevar = ccode,origin = "iso3n",destination = "country.name"),
               region=countrycode::countrycode(sourcevar = ccode,origin = "iso3n",destination = "region23")) %>% 
  select(ccode,country2,country3,everything())
```

```{r save}
save(df2,file = here::here("input/data/proc/study1_country.RData"))
sjPlot::view_df(df2,file =here::here("input/data/proc/codebook-s1.html") )
```

### 2025-07-29 procesamiento/04-prod_prep-supplementary-material.Rmd
data for reproducibility

if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4)  
# rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
data_corruption <- xlsx::read.xlsx("input/data/original/coruption-data/CPI2017_Full_DataSet-1801.xlsx",sheetIndex = 1)
data_corruption$year <- 2017
data_unionden <- haven::read_dta("input/data/original/union-density/union-density.dta") 
data_migration <- data.table::fread("input/data/original/migration/migrant-stock-share.csv")
data_unemployment <- data.table::fread("input/data/original/unemployment-world-bank/unemployment-rate.csv")
data_qog <- haven::read_dta("input/data/original/QOG/qog_std_ts_jan25.dta") 
data_vdem <- haven::read_dta("input/data/original/QOG/V-Dem-CY-Full+Others-v15.dta")

# data_socx <- data.table::fread("input/data/original/ocde-socx/ocde-socx.csv")
data_socx <- data.table::fread("input/data/original/ocde-socx/ocde-socspen.csv")

>>>>

if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4)  
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
data_corruption <- xlsx::read.xlsx("input/data/original/coruption-data/CPI2017_Full_DataSet-1801.xlsx",sheetIndex = 1)
data_corruption$year <- 2017
data_unionden <- haven::read_dta("input/data/original/union-density/union-density.dta") 
data_migration <- data.table::fread("input/data/original/migration/migrant-stock-share.csv")
data_unemployment <- data.table::fread("input/data/original/unemployment-world-bank/unemployment-rate.csv")
data_qog <- haven::read_dta("input/data/original/QOG/qog_std_ts_jan25.dta") 
data_vdem <- haven::read_dta("input/data/original/QOG/V-Dem-CY-Full+Others-v15.dta") 
data_socx <- data.table::fread("input/data/original/ocde-socx/ocde-socspen.csv")

### 2025-07-29 procesamiento/99-prod_analysis-descriptives.Rmd
data for reproducibility

>>>>

```{r}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif"))
options(scipen=999)
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,ggplot2,haven,ggrepel,gridExtra)  
rm(list=ls())

load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
df1$dclass3spo_V <- car::recode(df1$dclass3spo_V,"NA='No information (Missing, No partner)'",levels = c("Service Class (I+II)","Intermediate class (III+IV)","Working Class (V+VI+VII)","No information (Missing, No partner)"))

df1$employed <- ifelse(df1$MAINSTAT==1,yes = 1,no = 0)
df1$unemployed <- ifelse(df1$MAINSTAT==2,yes = 1,no = 0)

# The unemployment rate for an occupation j was computed as [(number of people unemployed in j)/
# (number of people unemployed in j + number of people employed in j)]*100

df_employment_country <- df1 %>% 
  group_by(country,country3,isco88r) %>% 
  summarise(unemployed=mean(unemployed),employed=mean(employed),isei08=mean(isei08)) %>% 
  mutate(unemrate=(unemployed/(unemployed+employed))*100)

df_employment_country %>% ungroup() %>% 
  select(unemployed,employed,isei08,unemrate) %>% 
  sjPlot::tab_corr()

dfreg <- df1 %>% dplyr::select(
  egal = egal2,
  dclass3res_V,
  dclass3spo_V,
  # homclass3_V_res,
  know_total,
  socialtrust,
  Q03pcm,
  edyears,
  female,
  agenum,
  religion,
  partner,
  union,
  workst,
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  gdppercapita,
  country2, country, country3,
) %>%
  # filter(oecd == "OECD") %>%
  filter(country2 != "SVN") %>% 
  na.omit() %>% 
    mutate(loggdppercapita=scale(gdppercapita/1000),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2, 
         # gini_disp=scale(gini_disp),
         )

### 2025-07-29 procesamiento/999-prod_prep-micro.Rmd
data for reproducibility

```{r}
if (!require("pacman")) install.packages("pacman") # instalar pacman
pacman::p_load(dplyr,sjmisc,GGally,sjlabelled,tidyverse,MLMusingR)
df1 <- haven::read_dta(here::here("input/data/original/ZA6980_v2-0-0.dta"))
dfTR <- haven::read_dta(here::here("input/data/original/ZA5521_v1-0-0.dta"))
dfES <- haven::read_dta(here::here("input/data/original/ZA5962_v1-0-0.dta"))
dfTR$weight <- 1
df1 <- df1 %>% rename(topbot=TOPBOT)
dfES <- dfES %>% rename(topbot=TOPBOT)
# fix problem with variable names
dfTR <- dfTR %>% rename(SEX=sex,
                        DEGREE=degree,
                        EDUCYRS=educyrs,
                        AGE=age,
                        MAINSTAT=mainstat,
                        UNION=union,
                        WEIGHT=weight,
                        URBRURAL=urbrural,
                        WORK=work,
                        HOMPOP=hompop,
                        EMPREL=emprel,
                        NSUP=nsup,
                        WRKSUP=wrksup,
                        SPEMPREL=spemprel,
                        SPWRKSUP=spwrksup,
                        ATTEND=attend, 
                        RELIGGRP=religgrp,
                        MARITAL=marital)

>>>>

```{r}
if (!require("pacman")) install.packages("pacman") # instalar pacman
pacman::p_load(dplyr,sjmisc,GGally,sjlabelled,tidyverse,MLMusingR)
df1 <- haven::read_dta(here::here("input/data/original/ZA6980_v2-0-0.dta"))
dfTR <- haven::read_dta(here::here("input/data/original/ZA5521_v1-0-0.dta"))
dfES <- haven::read_dta(here::here("input/data/original/ZA5962_v1-0-0.dta"))
dfTR$weight <- 1
df1 <- df1 %>% rename(topbot=TOPBOT)
dfES <- dfES %>% rename(topbot=TOPBOT)
# fix problem with variable names
dfTR <- dfTR %>% rename(SEX=sex,
                        DEGREE=degree,
                        EDUCYRS=educyrs,
                        AGE=age,
                        MAINSTAT=mainstat,
                        UNION=union,
                        WEIGHT=weight,
                        URBRURAL=urbrural,
                        WORK=work,
                        HOMPOP=hompop,
                        EMPREL=emprel,
                        NSUP=nsup,
                        WRKSUP=wrksup,
                        SPEMPREL=spemprel,
                        SPWRKSUP=spwrksup,
                        ATTEND=attend, 
                        RELIGGRP=religgrp,
                        MARITAL=marital,
                        WRKHRS=wrkhrs)
